FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-01-17 epss_marmo.tex

FITS is already able to propose standard formatting for some data products quite common by now in
planetary surface investigations.
In particular, Multi-Extension FITS (MEF) schema proposes an easy way
to store inhomogeneous digital information (reflectance, calibration data, vector table data, etc.)
in the same file each with relative metadata, as well as
multi-detector imagery (e.g. from HiRISE \citep{hirise}) or hyperspectral cubes
(e.g. from CRISM \citep{crism} or OMEGA \citep{omega}).
FITS has been already chosen to distribute, e.g.,
Hayabusa [10] and some of the Dawn [11] data.
To be efficiently used in planetary surface investigations, FITS metadata must be extended in order to
take into account the size of the reference body.
planetary body shapes and orientations are well defined
and standardized by the International Astronomical
Union (IAU) Working Group on Cartographic
Coordinates and Rotational Elements (WGCCRE).
This group reports triennially on the preferred rotation
rate, spin axis, prime meridian, and reference surface
for planets and satellites which helps ensure that
cartographic endeavors are effectively comparable [8].
In the framework of the VESPA [12, 13] component of the
Europlanet 2020 project an extension to FITS meta-
data (GeoFITS) has been proposed [14].

>>>>

FITS is already able to propose standard formatting for some data products quite
common by now in planetary surface investigations.
In particular, Multi-Extension FITS (MEF) schema proposes an easy way
to store inhomogeneous digital information (reflectance, calibration data, vector
table data, etc.) in the same file each with relative metadata, as well as
multi-detector imagery (e.g. from HiRISE \citep{hirise}) or hyperspectral cubes
(e.g. from CRISM \citep{crism} or OMEGA \citep{omega}).
FITS has been already chosen to distribute, e.g.,
Hayabusa\footnote{http://darts.isas.jaxa.jp/planet/project/hayabusa/index.html},
Akatsuki\footnote{http://darts.isas.jaxa.jp/planet/project/akatsuki/index.html}
and some of the Dawn\footnote{https://sbn.psi.edu/pds/archive/dawn.html} data.
To be efficiently used in planetary surface investigations, FITS metadata must
be extended in order to take into account the size of the reference body.
Planetary body shapes and orientations are well defined and standardized by the
International Astronomical Union (IAU) Working Group on Cartographic
Coordinates and Rotational Elements (WGCCRE).
This group reports triennially on the preferred rotation rate, spin axis, prime meridian,
and reference surface for planets and satellites which helps ensure that
cartographic endeavors are effectively comparable \citep{wgccre}.
In the framework of the VESPA [12, 13] component of the
Europlanet 2020 project an extension to FITS meta-
data (GeoFITS) has been proposed [14].
